---
title: "NFL Survival Pool Project"
author: "Jason Goldrosen"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: github_document
---
```{r setup, include = FALSE, cache = FALSE}  
knitr::opts_chunk$set(error = TRUE)  
```
## Background

There's a big NFL survival pool taking at work this year. Distracted by the NBA, Premier League, and all things Michigan sports, I haven't followed the NFL that closely over the last few years. 
In order to set my picks on auto-pilot, I'm an going to implement an Linear Integer Programming solution to maximize the probabilities of surviving throughout the season.  

## Previewing and cleaning the data
```{r preview data, warning = FALSE}
library(lpSolve)
library(ggplot2)


df <- read.csv("nfl_games_2019.csv") 
df$game_id <- c(1:nrow(df))
head(df)

```

```{r clean data}
df.1 <- df[,c(1,2,5,9,13)]
df.2 <- df[,c(1,2,6,9,13)]

names(df.1)[3:4] <- c("team","elo_prob")
names(df.2)[3:4] <- c("team","elo_prob")

df.2$elo_prob <- 1 - df.2$elo_prob
df.long <- rbind(df.1,df.2)
df.long <- df.long[order(df.long$game_id),]


df.long$date <- as.Date(df.long$date) 
df.long$dofw <- as.POSIXlt(df.long$date)$wday
df.long$date.2 <- df.long$date + (df.long$dofw == 4)*3 + (df.long$dofw == 1)*(-1) + (df.long$dofw == 6) + 1 
df.long$week <- as.integer((df.long$date.2-as.Date('2019-09-02'))/7)

```

## Implementing lpSolve

You can also embed plots, for example:

```{r lpsolve}
const <- matrix(rep(0,25088), 
                nrow=49, 
                ncol=512)
i <- 1
for (t in unique(df.long$team)){
const[i,] <- df.long$team == t
i <- i + 1
}

for (t in unique(df.long$week)){
  const[i,] <- df.long$week == t
  i <- i + 1
}
obj <- df.long$elo_prob
direction <- c(rep("<=",length(unique(df.long$team))),rep("=",length(unique(df.long$week))))
rhs <- rep(1,nrow(const))

test <- lp(direction = "max", 
           objective.in = obj, 
           const.mat = const, 
           const.dir = direction, 
           const.rhs = rhs)

solns <- df.long[test$solution!=0,c("week","team","elo_prob")]
#colnames(solns) <- c("Week", "Team", "Pr(Win)")

knitr::kable(solns, row.names=FALSE, digits = 4)

```


```{r plots}
solns$prob <- cumprod(solns$elo_prob)

ggplot(solns, aes(x=week,y=prob)) + 
  geom_point() + 
  geom_line() + 
  labs(title="Cumulative Survival Probability",x ="NFL Game Week", y = "")

```
